---
title: "Bulls_Player_Metric"
author: "Sara G"
date: "28/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Opening libraries

```{r librarys}
library(tidyverse)
library(dplyr)
library(broom)
```

# Reading in and tidying Bulls data files

```{r read_data, message=TRUE} 
sal <- read_csv("data/raw/2018-19_nba_player-salaries.csv")
stat_pl <- read_csv("data/raw/2018-19_nba_player_statistics.csv")
stat_tm1 <- read_csv("data/raw/2018-19_nba_team_statistics_1.csv")
stat_tm2 <- read_csv("data/raw/2018-19_nba_team_statistics_2.csv")
pay <- read_csv("data/raw/2019-20_nba_team-payroll.csv")
```

# Tidying data

## Binding data from two data frames
Binding variables from team 1 data and team 2 data

```{r binding}
df_team <- left_join(x = stat_tm2, y = stat_tm1 [-c(1,3, 23:25)],
                     by = "Team")
```

## renaming variables: 
Making sure variables do not include a %, replace with 'p'. And added a 'x' in front of number variables. Change '/' to _per_ .

```{r renaming}
df_team <- df_team %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%', x3PAr = "3PAr", TSp = "TS%", eFGp = "eFG%", TOVp = "TOV%", ORBp = "ORB%", FT_per_FGA = "FT/FGA", DRBp = "DRB%")
```


## Normalising key metrics

We want normalise the key metrics to per game


## Normalising variables per minuted played minutes played
There are 5 players playing roughly 48mins per game. Therefore there are ~240mins per game. WE want normalise these varaibles for the minutes played. 

```{r Normalising metrics_pmp}
df_team <- df_team %>%
  mutate (PTS_pmp = PTS / MP , 
          AST_pmp = AST / MP,
          FG_pmp = FGA / MP, 
          x3P_pmp = x3P / MP ,
          x2P_pmp = x2P / MP ,
          FT_pmp = FTA / MP,
          STL_pmp = STL / MP ,
          BLK_pmp = BLK / MP,
          TOV_pmp = TOV / MP,
          ORB_pmp = ORB / MP,
          DRB_pmp = DRB / MP,
          TRB_pmp = TRB / MP,
          PF_pmp = PF / MP)
        
```

##Check structure of new data set

```{r Str_df_team}
str(df_team)
```

# Team metric for scoring points per minute 

```{r new lm wmp}
fit_mp <- lm(PTS_pmp ~   x3P_pmp + x2P_pmp + FT_pmp, data = df_team)
tidy(fit_mp, conf.int = TRUE)
```

## Creating player metric

### Tidying player data

```{r renaming_player}
df_player <- stat_pl %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%', eFGp = "eFG%")
```


## Check struture of player stats data

```{r str player}
str(df_player)
```

```{r head player }
head(df_player)
```

```{r tail playe}
tail(df_player)

```

# There are multiple players players on each row. They need to be summarised in one row per player, only chosing the varaibles that we want to look at. 

```{r summaries players}
df_play_sum <- df_player %>%
   select(player_name, G, GS, MP, FG, x3P, x2P, FT, PTS) %>%
  group_by(player_name) %>%
  summarise(MP = sum(MP),
            G = sum(G),
            GS = sum(G),
            FG = sum(FG), 
            x3P = sum(x3P),
            x2P = sum(x2P),
            FT = sum(FT), 
            PTS = sum(PTS))
```


```{r head player summary}
head(df_play_sum)
```

```{r tail player summary}
tail(df_play_sum)
```

```{r filter playing position}
pos <- df_player %>%
  filter(G > 0) %>%
  group_by(player_name, Pos) %>%
  select(player_name, Pos, G, GS)

```

## Checking all the positions

```{r frequency distribution}
table(pos$Pos)

```


## Eliminating positions that are more than one person


```{r }
pos <- pos %>%
  filter(Pos != "C-PF", Pos != "PF-C", Pos != "PF-SF", Pos != "SF-SG", Pos != "SG-PF", Pos != "SG-SF" )

```


## Rechecking the positions
```{r frequency distribution}
table(pos$Pos)

```


#Averaging out minutes played per game, more player (mp_per_game) to create one number from the team data for the player data

```{r normalising players min played}
mp_per_game <- df_team %>%
  group_by(Team) %>%
  summarise(mp_per_game = MP/G/5) %>%
  pull(mp_per_game) %>%
  mean

mp_per_game
```


## Creating new variables: Points per minute per player for each of the model metrics

```{r Player ave MP}
df_play_sum <- df_play_sum %>%
 group_by(player_name) %>%
  mutate(PTS_pmp = PTS / MP,
            x3P_pmp = x3P / MP,
            x2P_pmp = x2P / MP,
            FT_pmp = FT / MP)
```

## normalising using the minutes played per game


#Identifying how all the below points were achieved, with the normalised per minutes played

```{r New PTS statistics}
player <- df_play_sum %>%
 summarise(MPn = sum(MP)/mp_per_game,
            x3P_pmp = sum(x3P_pmp)/ MPn,
            x2P_pmp = sum(x2P_pmp)/MPn,
            FT_pmp = sum(FT_pmp) / MPn,
           G = sum(G)) %>%
  filter(G > 20) %>%
  select(-c(MPn))

head(player)

```

## Expected points per minuted played metric

```{r player model}
player <- mutate(player, exp_PTS_ppl = predict(fit_mp, newdata = player))

```

#Expected points per player metric visualisation

```{r noraml distribution}
player %>%
  ggplot(aes(x = exp_PTS_ppl)) +
  geom_histogram(binwidth = 0.01, colour = "black", fill = "blue")

```

## add in salary

```{r salary}
player <- sal %>%
  select(player_name, salary) %>%
  right_join(player, by = "player_name")

```


## FIltering out the multiple positions per player

```{r filter playing position}
pos <- pos %>%
  filter(G > 0) %>%
  group_by(player_name, Pos) %>%
  summarise(G = sum(G)) %>%
  top_n(1,abs(G))

```


## add in player positions

```{r add positions}
player <- pos %>%
  select(player_name, Pos) %>%
  right_join(player, by = "player_name")
```  

#Checking for NAs

```{r checking nas}
sum(is.na (player))
```

#filtering out NAs

```{r nas}  
player <- player %>%
  filter(!is.na(Pos)) %>%
  filter(!is.na(salary))
        
```



# checking for NAs again

```{r checking nas}
sum(is.na (player))
```


#Creating Player rating data frame with ideal players at the top

```{r selecting players}
player_choices <- player %>%
  select(player_name, Pos, salary, exp_PTS_ppl) %>%
  arrange(desc(exp_PTS_ppl), salary)

player_choices

```

```{r Selecting players graph}
gg_play_sal_vs_Expt_PTS <- player_choices %>%
  ggplot(aes(x = salary / 1000, y = (exp_PTS_ppl *48*5), color = Pos)) +
  geom_point() +
  xlab("Salary (Thousands)") 
gg_play_sal_vs_Expt_PTS

```

```{r final graph of exp points vs salary}
gg_play_sal_vs_Expt_PTS <- gg_play_sal_vs_Expt_PTS +
  ggtitle("Expect points per game per player by salary") +
  ylab("Expected points per player")

gg_play_sal_vs_Expt_PTS
``` 

```{r save final player graph}
ggsave("figs/Exp_points_per_player_by_salary.pdf")
```





```{r Player ave MP}
player <- df_play_sum %>%
 group_by(player_name) %>%
  mutate(PTS_n = PTS_pmp * mp_per_game,
            x3P_n = x3P_pmp * mp_per_game,
            x2P_n = x2P_pmp * mp_per_game,
            FT_n = FT_pmp * mp_per_game) %>%
  filter(PTS_n > 10) %>%
  select(c(player_name, Pos, PTS_n, x3P_n, x2P_n, FT_n, x3P_pmp, x2P_pmp, FT_pmp, PTS_pmp))
```


```{r Normalising metrics_pg}
df_team <- df_team %>%
  mutate (PTS_per_game = PTS / G,
          WINp = W/G,
          AST_per_game = AST / G,
          FG_per_game = FGA / G, 
          x3P_per_game = x3P / G,
          x2P_per_game = x2PA / G,
          FT_per_game = FTA / G,
          STL_per_game = STL / G,
          BLK_per_game = BLK / G,
          TOV_per_game = TOV / G,
          ORB_per_game = ORB / G,
          DRB_per_game = DRB / G,
          TRB_per_game = TRB / G,
          PF_per_game = PF / G)
```