---
title: "Bulls_Player_Metric"
author: "Sara G"
date: "28/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Opening libraries

```{r librarys}
library(tidyverse)
library(dplyr)
library(broom)
```

# Reading in and tidying Bulls data files

```{r read_data, message=TRUE} 
sal <- read_csv("data/raw/2018-19_nba_player-salaries.csv")
stat_pl <- read_csv("data/raw/2018-19_nba_player_statistics.csv")
stat_tm1 <- read_csv("data/raw/2018-19_nba_team_statistics_1.csv")
stat_tm2 <- read_csv("data/raw/2018-19_nba_team_statistics_2.csv")
pay <- read_csv("data/raw/2019-20_nba_team-payroll.csv")
```

# Tidying data

## Binding data from two data frames
Binding variables from team 1 data and team 2 data

```{r binding}
df_team <- left_join(x = stat_tm2, y = stat_tm1 [-c(1,3, 23:25)],
                     by = "Team")
```

## renaming variables: 
Making sure variables do not include a %, replace with 'p'. And added a 'x' in front of number variables. Change '/' to _per_ .

```{r renaming}
df_team <- df_team %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%', x3PAr = "3PAr", TSp = "TS%", eFGp = "eFG%", TOVp = "TOV%", ORBp = "ORB%", FT_per_FGA = "FT/FGA", DRBp = "DRB%")
```


## Normalising key metrics

We want normalise the key metrics to per game

```{r Normalising metrics_pg}
df_team <- df_team %>%
  mutate (PTS_per_game = PTS / G,
          WINp = W/G,
          AST_per_game = AST / G,
          FG_per_game = FGA / G, 
          x3P_per_game = x3P / G,
          x2P_per_game = x2PA / G,
          FT_per_game = FTA / G,
          STL_per_game = STL / G,
          BLK_per_game = BLK / G,
          TOV_per_game = TOV / G,
          ORB_per_game = ORB / G,
          DRB_per_game = DRB / G,
          TRB_per_game = TRB / G,
          PF_per_game = PF / G)
```

## Normalising variables per minuted played minutes played
There are 5 players playing roughly 48mins per game. Therefore there are ~240mins per game. WE want normalise these varaibles for the minutes played. 

```{r Normalising metrics_pmp}
df_team <- df_team %>%
  mutate (PTS_pmp = PTS / MP , 
          AST_pmp = AST / MP,
          FG_pmp = FGA / MP, 
          x3P_pmp = x3P / MP ,
          x2P_pmp = x2P / MP ,
          FT_pmp = FTA / MP,
          STL_pmp = STL / MP ,
          BLK_pmp = BLK / MP,
          TOV_pmp = TOV / MP,
          ORB_pmp = ORB / MP,
          DRB_pmp = DRB / MP,
          TRB_pmp = TRB / MP,
          PF_pmp = PF / MP)
        
```

##Check structure of new data set

```{r Str_df_team}
str(df_team)
```

# Team metric for scoring points per minute 

```{r new lm wmp}
fit_mp <- lm(PTS_pmp ~   x3P_pmp + x2P_pmp + FT_pmp, data = df_team)
tidy(fit_mp, conf.int = TRUE)
```

## Creating player metric

### Tidying player data

```{r renaming_player}
df_player <- stat_pl %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%', eFGp = "eFG%")
```


```{r normalising players min played}
mp_per_game <- df_team %>%
  group_by(Team) %>%
  summarise(mp_per_game = MP/G /5) %>%
  pull(mp_per_game) %>%
  mean

mp_per_game
```


## Points per minute per player

```{r Player ave MP}
df_player <- df_player %>%
 group_by(player_name) %>%
  mutate(PTS_pmp = PTS / MP,
            x3P_pmp = x3P / MP,
            x2P_pmp = x2P / MP,
            FT_pmp = FTA / MP)
```



```{r Player ave MP}
player <- df_player %>%
 group_by(player_name) %>%
  mutate(PTS_n = PTS_pmp * mp_per_game,
            x3P_n = x3P_pmp * mp_per_game,
            x2P_n = x2P_pmp * mp_per_game,
            FT_n = FT_pmp * mp_per_game) %>%
  filter(PTS_n > 10) %>%
  select(c(player_name, Pos, PTS_n, x3P_n, x2P_n, FT_n, x3P_pmp, x2P_pmp, FT_pmp, PTS_pmp))
```


```{r New PTS statistics}
player <- df_player %>%
 summarise(MPn = sum(MP)/mp_per_game,
            x3P_pmp = sum(x3P_pmp)/ MPn,
            x2P_pmp = sum(x2P_pmp)/MPn,
            FT_pmp = sum(FT_pmp) / MPn,
           G = sum(G)) %>%
  filter(G > 20) %>%
  select(-c(MPn))

head(player)

```


```{r player model}
player <- mutate(player, exp_PTS_ppl = predict(fit_mp, newdata = player))

```

#Expected points per player metric

```{r noraml distribution}
player %>%
  ggplot(aes(x = exp_PTS_ppl)) +
  geom_histogram(binwidth = 0.01, colour = "black", fill = "blue")

```

## add in salary

```{r salary}
player <- sal %>%
  select(player_name, salary) %>%
  right_join(player, by = "player_name")

```


## add in player positions



```{r filter playing position}
pos <- df_player %>%
  filter(G > 20) %>%
  group_by(player_name, Pos) %>%
  summarise(G = sum(G)) %>%
  top_n(1,abs(G))
 

  

```

```{r add positions}
player <- pos %>%
  select(player_name, Pos) %>%
  right_join(player, by = "player_name") %>%
```  
  
```{r nas}  
player <- player %>%
  filter(!is.na(Pos)) %>%
  filter(!is.na(salary))
           

```


```{r selecting players}
player <- player %>%
  select(player_name, Pos, salary, exp_PTS_ppl) %>%
  arrange(desc(exp_PTS_ppl), salary)

```


