---
title: "Analysis_bulls"
author: "Sara G"
date: "27/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Opening libraries

```{r librarys}
library(tidyverse)
library(dplyr)
library(broom)
```


# Reading in and tidying Bulls data files

```{r read_data, message=TRUE} 
sal <- read_csv("data/raw/2018-19_nba_player-salaries.csv")
stat_pl <- read_csv("data/raw/2018-19_nba_player_statistics.csv")
stat_tm1 <- read_csv("data/raw/2018-19_nba_team_statistics_1.csv")
stat_tm2 <- read_csv("data/raw/2018-19_nba_team_statistics_2.csv")
pay <- read_csv("data/raw/2019-20_nba_team-payroll.csv")
```

# Tidying data

## Binding data from two data frames
Binding variables from team 1 data and team 2 data

```{r binding}
df_team <- left_join(x = stat_tm2, y = stat_tm1 [-c(1,3, 23:25)],
                     by = "Team")
```

## renaming variables: 
Making sure variables do not include a %, replace with 'p'. And added a 'x' in front of number variables. Change '/' to _per_ .

```{r renaming}
df_team <- df_team %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%', x3PAr = "3PAr", TSp = "TS%", eFGp = "eFG%", TOVp = "TOV%", ORBp = "ORB%", FT_per_FGA = "FT/FGA", DRBp = "DRB%")
```

## Normalising key metrics

We want normalise the key metrics to per game

```{r Normalising metrics_pg}
df_team <- df_team %>%
  mutate (PTS_per_game = PTS / G,
          WINp = W/G,
          AST_per_game = AST / G,
          FG_per_game = FGA / G, 
          x3P_per_game = x3P / G,
          x2P_per_game = x2PA / G,
          FT_per_game = FTA / G,
          STL_per_game = STL / G,
          BLK_per_game = BLK / G,
          TOV_per_game = TOV / G,
          ORB_per_game = ORB / G,
          DRB_per_game = DRB / G,
          TRB_per_game = TRB / G,
          PF_per_game = PF / G)
```

## Normalising variables per minuted played minutes played
There are 5 players playing roughly 48mins per game. Therefore there are ~240mins per game. WE want normalise these varaibles for the minutes played. 

```{r Normalising metrics_pmp}
df_team <- df_team %>%
  mutate (PTS_pmp = PTS / MP , 
          AST_pmp = AST / MP,
          FG_pmp = FGA / MP, 
          x3P_pmp = x3P / MP ,
          x2P_pmp = x2P / MP ,
          FT_pmp = FTA / MP,
          STL_pmp = STL / MP ,
          BLK_pmp = BLK / MP,
          TOV_pmp = TOV / MP,
          ORB_pmp = ORB / MP,
          DRB_pmp = DRB / MP,
          TRB_pmp = TRB / MP,
          PF_pmp = PF / MP)
        
```

##Check structure of new data set

```{r Str_df_team}
str(df_team)
```



## Multiple Regression

```{r MLR per game}
lm_teams_pg <- lm(PTS_per_game ~ AST_per_game + FG_per_game + x3P_per_game + x2P_per_game + FT_per_game + STL_per_game + BLK_per_game + TOV_per_game + DRB_per_game + ORB_per_game, data = df_team)
tidy(lm_teams_pg, conf.int = TRUE)
```

```{r summary}
summary(lm_teams_pg)
```


## Checking assumptions for MLR
Checking assumptions for linear model (lm_teams_pg)

# CHoosing variables that have a relationship.

```{r MLR per game}
lm_teams_pg <- lm(PTS_per_game ~ AST_per_game + FG_per_game + x3P_per_game + x2P_per_game + FT_per_game + STL_per_game + BLK_per_game + TOV_per_game + DRB_per_game + ORB_per_game, data = df_team)
tidy(lm_teams_pg, conf.int = TRUE)
```





1, Response variables should be continuous. Yes this is met
2. There are wo or more explanatory variables that are continuous or categorical Yes all of variables are continous
3. Independence

```{r Assumpt_independence}
car::durbinWatsonTest(lm_teams_pg)
```

4. Linearity

```{r linearity}
car::avPlots(lm_teams_pg)
```

Update variables to include only those with a relationship

```{r new lm}
fit_pg <- lm(PTS_per_game ~ AST_per_game + x3P_per_game + x2P_per_game + FT_per_game, data = df_team)
tidy(fit_pg, conf.int = TRUE)
```

3. Independence

```{r Assumpt_independence}
car::durbinWatsonTest(fit_pg)
```

4. Linearity

```{r linearity}
car::avPlots(fit_pg)
```

5. Outliers, Influential or high leverage points

```{r outliers}
std_res <- rstandard(fit_pg)
points <- 1:length(std_res)

ggplot(data = NULL, aes(x = points, y = std_res)) +
  geom_point() +
  ylim(c(-4,4)) +
  geom_hline(yintercept = c(-3,3), colour = "red", linetype = "dashed")
```

```{r outliers labeled}
res_labels <- if_else(abs(std_res) >= 2, paste(points), "")
                    
ggplot(data = NULL, aes(x = points, y = std_res, label = res_labels)) +
  geom_point() +
  ylim(c(-4,4)) +
  geom_text(nudge_x = 0.5 ) +
  geom_hline(yintercept = c(-3,3), colour = "red", linetype = "dashed")
```

## Leverage points

```{r hats}
hats <- hatvalues(fit_pg)
hats_labels <- if_else(hats >= 0.2, paste(points), "")

ggplot(data = NULL, aes(x = points, y = hats)) +
  geom_point() +
  geom_text(aes(label = hats_labels), nudge_y = 0.02)
```

# There are no hat values that are greater than 1 but should investigate the hat values that are greater than 0.2 as they are above the rest

## Influential points

```{r Cook}
cook <- cooks.distance(fit_pg)
cook_labels <- if_else(cook >= 0.1, paste(points), "")

ggplot(data = NULL, aes(x = points, y = cook)) +
  geom_point() +
  geom_text(aes(label = cook_labels), nudge_y = 0.02)

```

# There are no points that are above 1, but we will look at points above 0.1

We will create a new df without the high influence points

```{r new df with no outliers}
outliers <- c(9, 11)
filtered_df <- df_team %>%
  filter(!Rk %in% outliers)

```

```{r new lm with filtered_df}
fit2 <- lm(PTS_per_game ~ AST_per_game + x3P_per_game + x2P_per_game + FT_per_game, data = filtered_df)

tidy(fit2, conf.int = TRUE)
```

```{r summary filtered_df}
summary(fit2)
```

# by filtering out these variables there is minimal change to the slope or the intercept or the R-square value (from 0.89 to 0.88).
# we don't want to take out these factors as they can introduce variability to the model and they make minimal change so we will leave them in. 

6. Homoscedasticity

```{r homoscedasticity}
res <- residuals(fit_pg)
fitted <- predic(fit_pg)

ggplot(data = NULL, aes(x = fitted, y = res)) +
  geom_point(colour = "dodgerblue") +
  geom_hline(yintercept = 0, colour = "red", linetype = "dashed")

```












