---
title: "Bulls_assign_Read_tidy_Explore"
author: "Sara G"
date: "25/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Opening libraries

```{r librarys}
library(tidyverse)
library(dplyr)
library(broom)
```

# Reading in and tidying Bulls data files

```{r read_data, message=TRUE} 
sal <- read_csv("data/raw/2018-19_nba_player-salaries.csv")
stat_pl <- read_csv("data/raw/2018-19_nba_player_statistics.csv")
stat_tm1 <- read_csv("data/raw/2018-19_nba_team_statistics_1.csv")
stat_tm2 <- read_csv("data/raw/2018-19_nba_team_statistics_2.csv")
pay <- read_csv("data/raw/2019-20_nba_team-payroll.csv")
```

```{r str_team2}
str(stat_tm2)
```

## Identify average minutes per game

We want to find how many of different variables occurred per minuted played. 

Therefore we want to identify the average minutes played per game

```{r ave minutes per game}
df_team <- stat_tm2 %>%
  mutate(AMPG = MP / G)
```

## renaming variables: 
Making sure variables do not include a %, replace with 'p'. And added a 'x' in front of number variables. 

```{r renaming}
df_team <- df_team %>%
  rename( x3P = '3P', x3PA = '3PA', x3Pp = '3P%', x2P = '2P', x2PA = '2PA', x2Pp = '2P%', FGp = 'FG%', FTp = 'FT%')
```

## Normalising key metrics

We want normalise the key metrics to per game

```{r Normalising metrics}
df_team <- df_team %>%
  mutate (PTS_per_game = PTS / G, 
          AST_per_game = AST / G,
          FGA_per_game = FGA / G, 
          x3PA_per_game = x3P / G,
          x2PA_per_game = x2PA / G,
          FTA_per_game = FTA / G,
          STL_per_game = STL / G,
          BLK_per_game = BLK / G,
          TOV_per_game = TOV / G)
```

## Normalising variables per minuted played per game
There are 5 players playing roughly 48mins per game. Therefore there are ~240mins pwe game. WE want normalise these varaibles for the minutes played. 

```{r Normalising metrics}
df_team <- df_team %>%
  mutate (PTS_pmp = PTS / MP, 
          AST_pmp = AST / MP,
          FGA_pmp = FGA / MP, 
          x3PA_pmp = x3P / MP ,
          x2PA_pmp = x2PA / MP ,
          FTA_pmp = FTA / MP,
          STL_pmp = STL / MP ,
          BLK_pmp = BLK / MP,
          TOV_pmp = TOV / MP)
```

## Explore Data

Comparing points per game to other team variables
See if they correlate with points

# Points vs Assists


```{r points vs Assists per minute played}
ggplot(df_team, aes(AST_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm")
```

```{r points vs Assists per game}
ggplot(df_team, aes(AST_per_game, y = PTS_per_game)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.5, nudge_x = 0.5)

```

# Points vs FGA


```{r points vs Field Goals per minute played}
ggplot(df_team, aes(FGA_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.005, nudge_x = 0.005)

```

```{r points vs Field Goals Attempts per game}
ggplot(df_team, aes(x = FGA_per_game, y = PTS_per_game)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.5, nudge_x = 0.5)

```

# Points vs x3PA


```{r points vs x3PA per minute played}
ggplot(df_team, aes(x3PA_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.005, nudge_x = 0.0005)

```

```{r points vs x3PA Attempts per game}
ggplot(df_team, aes(x = x3PA_per_game, y = PTS_per_game)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm")
 
```


# Points vs Steals

```{r points vs Steals per minute played}
ggplot(df_team, aes(STL_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.005, nudge_x = 0.001)

```


# Points vs Blocks

```{r points vs Blocks per minute played}
ggplot(df_team, aes(BLK_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.003, nudge_x = 0.0005)

```

```{r points vs Blocks per game}
ggplot(df_team, aes(BLK_per_game, y = PTS_per_game)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.5, nudge_x = 0.25)

```


# Points vs Free Throws

```{r points vs Free Throws Attempts per minute played}
ggplot(df_team, aes(FTA_pmp, y = PTS_pmp)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.002, nudge_x = 0.002, size = 3)

```


```{r points vs Free Throw Attempts per game}
ggplot(df_team, aes(FTA_per_game, y = PTS_per_game)) +
  geom_point(alpha = 0.5, colour = "purple") +
  geom_smooth(method = "lm") +
  geom_text(aes(label = Team), nudge_y = 0.25, nudge_x = 0.25)

```

## Checking for confounding variables

```{r confounding}
pairs(formula = ~ AST_pmp + FGA_pmp + x3PA_pmp + x2PA_pmp + FTA_pmp + STL_pmp + BLK_pmp + TOV_pmp, data = df_team)
```

## Multiple Regression

```{r MLR per game}
lm_teams_pg <- lm(PTS_per_game ~ AST_per_game + FGA_per_game + x3PA_per_game + x2PA_per_game + FTA_per_game + STL_per_game + BLK_per_game + TOV_per_game, data = df_team)
tidy(lm_teams_pg, conf.int = TRUE)
```

```{r summary per game}
summary(lm_teams_pg)
```


```{r MLR per minute played}
lm_teams_pmp <- lm(PTS_pmp ~ AST_pmp + FGA_pmp + x3PA_pmp + x2PA_pmp + FTA_pmp + STL_pmp + BLK_pmp + TOV_pmp, data = df_team)
tidy(lm_teams_pmp, conf.int = TRUE)
```

```{r summary per minute played}
summary(lm_teams_pmp)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
